---
title: "AI Trust"
output: html_notebook
---

# R environment

```{r}
# R environment
R.Version()
```

# Install necessary packages

```{r}
 # install.packages(mosaic)
 # install.packages(supernova)
 # install.packages(Lock5withR)
 # install.packages(tidyverse)
 # install.packages(magrittr)
 # install.packages(gridExtra)
 # install.packages(stringr)
 # install.packages(dplyr)
 # install.packages(lme4)
 # install.packages(broom)
 # install.packages(emmeans)
```

# Download necessary packages

```{r}
suppressPackageStartupMessages({
   library(mosaic)
   library(supernova)
   library(Lock5withR)
   library(tidyverse)
   library(magrittr)
   library(gridExtra)
   library(stringr)
   library(dplyr)
   library(lme4)
   library(broom)
   library(emmeans)
    })

```

# Load Data

```{r}
Raw_AItrust <- read_csv("Excel_Cleaned.csv")
```

# Data Cleaning

### Rename Columns

```{r}
Clean_AItrust <- Raw_AItrust %>% 
  rename( 
        'Subject' = 'ResponseId',
        'BaselineTrust'='Baseline_1', #Component of Dependent Variable
        'FinalTrust'='FinalTrust_1', #Component of Dependent Variable
        "DurationSec" = 'Duration (in seconds)',
        #The following columns are data for exploratory analysis that were not included in the paper
        "TrustQ1_1" = 'TrustQ1_1...27' , 
         "FeelQ1_1" = 'FeelQ1_1...28',
         "TrustQ2_1" = 'TrustQ2_1...30' ,
         "FeelQ2_1" = 'FeelQ2_1...31' ,
        "TrustQ3_1" = 'TrustQ3_1...33',
         "FeelQ3_1" = 'FeelQ3_1...34' ,
        #####
         "TrustQ1_2" = 'TrustQ1_1...36',
         "FeelQ1_2" = 'FeelQ2_1...37' ,
         "TrustQ2_2" = 'TrustQ2_1...39',
         "FeelQ2_2" = 'FeelQ2_1...40',
         "TrustQ3_2" = 'TrustQ3_1...42',
         "FeelQ3_2" = 'FeelQ3_1...43',
        #####
         "TrustQ1_3" = 'TrustQ1_1...45',
         "FeelQ1_3" = 'FeelQ1_1...46',
         "TrustQ2_3" = 'TrustQ2_1...48',
         "FeelQ2_3" = 'FeelQ3_1...49' ,
         "TrustQ3_3" =  'TrustQ3_1...51',
         "FeelQ3_3" = 'FeelQ3_1...52',
        #####
         "TrustQ1_4" = 'TrustQ1_1...54',
         "FeelQ1_4" = 'FeelQ1_1...55',
         "TrustQ2_4" = 'TrustQ2_1...57',
         "FeelQ2_4" = 'FeelQ2_1...58',
         "TrustQ3_4" = 'TrustQ3_1...60',
         "FeelQ3_4" = 'FeelQ3_1...61' 
        )
```

### Combined columns

This chunk of code is for exploratory analysis that were not included in the paper

```{r}
# Combined trust ratings for the first ChatGPT response in each condition into a single column called TrustQ1
Clean_AItrust$TrustQ1 <- coalesce(Clean_AItrust$TrustQ1_1, Clean_AItrust$TrustQ1_2, Clean_AItrust$TrustQ1_3,Clean_AItrust$TrustQ1_4) 

# Combined trust ratings for the second ChatGPT response in each condition into a single column called TrustQ2
Clean_AItrust$TrustQ2 <- coalesce(Clean_AItrust$TrustQ2_1, Clean_AItrust$TrustQ2_2, Clean_AItrust$TrustQ2_3,Clean_AItrust$TrustQ2_4) 

# Combined trust ratings for the third ChatGPT response in each condition into a single column called TrustQ3
Clean_AItrust$TrustQ3 <- coalesce(Clean_AItrust$TrustQ3_1, Clean_AItrust$TrustQ3_2, Clean_AItrust$TrustQ3_3,Clean_AItrust$TrustQ3_4) 

# Combined feeling ratings for the first ChatGPT response in each condition into a single column called FeelingQ1
Clean_AItrust$FeelingQ1 <- coalesce(Clean_AItrust$FeelQ1_1, Clean_AItrust$FeelQ1_2, Clean_AItrust$FeelQ1_3,Clean_AItrust$FeelQ1_4) 

# Combined feeling ratings for the second ChatGPT response in each condition into a single column called FeelingQ2
Clean_AItrust$FeelingQ2 <- coalesce(Clean_AItrust$FeelQ2_1, Clean_AItrust$FeelQ2_2, Clean_AItrust$FeelQ2_3,Clean_AItrust$FeelQ2_4) 

# Combined feeling ratings for the third ChatGPT response in each condition into a single column called FeelingQ3
Clean_AItrust$FeelingQ3 <- coalesce(Clean_AItrust$FeelQ3_1, Clean_AItrust$FeelQ3_2, Clean_AItrust$FeelQ3_3,Clean_AItrust$FeelQ3_4) 

```

### Select analysis relevant columns

```{r}
Clean_AItrust <- Clean_AItrust %>% 
    select(
      'College', #Participants' academic information
        'Status', #Show Survey Preview Response and Real Response 
        'Subject', #Participant ID
       'Progress', #Survey Progress
        "Integrity", #The Integrity and Attention Check Questions 
        'BaselineTrust', #Participants Trust in ChatGPT before the experiment
        "FinalTrust", #Participants Trust in ChatGPT at the end of experiment
        "Version", #Participants' Condition (Independent Variables:Framing Message and ChatGPT response accuracy)
        
      # The following variables are for exploratory analysis that's not included in the paper
       'TrustQ1', #Participants Trust in the first ChatGPT's response
        'TrustQ2', #Participants Trust in the second ChatGPT's response  
        'TrustQ3', #Participants Trust in the third ChatGPT's response  
        "FeelingQ1", #Participants Feeling towards the first ChatGPT's response  
         "FeelingQ2", #Participants Feeling towards the second ChatGPT's response 
         "FeelingQ3" #Participants Feeling towards the third ChatGPT's response  
        )
```

### Remove non-response cases

```{r}
Clean_AItrust = Clean_AItrust %>%
slice(-(1:2)) #getting rid of rows one and two
```

### Remove ineligible responses

1.  Is Survey preview response
2.  Failed Integrity and attention check

```{r}
Clean_AItrust = Clean_AItrust %>%
    filter(Status != 'Survey Preview' & Integrity != 'No')

## remove these columns afterwards
Clean_AItrust = Clean_AItrust %>%
    subset(select = -c(Status, Integrity))
```

### Create column for Independent Variables

```{r}
#Info of Framing Message (Negative or Neutral) in Version column store in a new column, FramingM
Clean_AItrust$FramingM <- str_extract(Clean_AItrust$Version, "Negative|Neutral") 

#Info of ChatGPT Response Accuracy (High or Low) in Version column store in a new column, Accuracy
Clean_AItrust$Accuracy <- str_extract(Clean_AItrust$Version, "High|Low")

# remove the version column afterwards
Clean_AItrust = Clean_AItrust %>%
    subset(select = -c(Version))
```

### Recode Data Value

```{r}
#The original data value Neutral in the FramingM column recode to No Messaging
Clean_AItrust <- Clean_AItrust %>% 
    mutate(FramingM = case_when(FramingM == "Neutral" ~ "No Messaging",TRUE ~ FramingM)) 
```

### Convert data types for IV(factors) and DV(integer)

```{r}
# Factors
Clean_AItrust <- Clean_AItrust %>% 
 mutate(
        FramingM = factor(FramingM, levels = c('No Messaging', 'Negative')),
        Accuracy = factor(Accuracy, levels = c('High', 'Low'))
        )

# DV: integer
Clean_AItrust <- Clean_AItrust %>% 
  mutate( 
          BaselineTrust = as.integer(BaselineTrust),
          FinalTrust = as.integer(FinalTrust),
        #The following code are for exploratory analysis that were not included
         TrustQ1 = as.integer(TrustQ1), 
         TrustQ2 = as.integer(TrustQ2),
         TrustQ3 = as.integer(TrustQ3),
         FeelingQ1 = as.integer(FeelingQ1),
         FeelingQ2 = as.integer(FeelingQ2),
         FeelingQ3 = as.integer(FeelingQ3)
         )
```

### Create Dependent Variable column

```{r}
Clean_AItrust <- Clean_AItrust %>% 
  mutate(
        TrustDiff = FinalTrust - BaselineTrust, #Dependent Variable: Participants' Change in Trust
        #The following code are for exploratory analysis that were not included
        Trust3Mean = ((TrustQ1 + TrustQ2 + TrustQ3)/3),  
        Feeling3Mean = ((FeelingQ1 + FeelingQ2 + FeelingQ3)/3) 
        ) 
```

# Data Visualization

### For Paper

```{r}
Clean_AItrust %>%
    ggplot(aes(y = TrustDiff, x = FramingM, color = Accuracy, width = 20, height = 50))+
    stat_summary(
        fun = mean, 
        geom = "point",
        size = 4) +
    stat_summary(
        fun = mean, 
        geom = "line", 
        size = 1, 
        aes(group = Accuracy)) +
    stat_summary(
        fun.data = mean_se, 
        geom = "errorbar", 
        width = 0.2, 
        size = 1)+
    labs(
         x = "Framing Message", 
         y = "Change in Trust in ChatGPT") +
    theme_bw( ) + # apply white background, black border to the plot
    theme(
          axis.title.x  = element_text(size = 12, face = "bold", margin = margin(t = 10),family = "Times"), #set the size, boldness, top margin and style of the x axis title 
          axis.title.y  = element_text(size = 12, face = "bold",margin = margin(r = 10),family = "Times"), #set the size, boldness, right margin and style of the y axis title 
          axis.text = element_text(size = 12, color = "black",family = "Times"), #set the size, color, and style of both axis texts
          legend.title = element_text(size = 12, face = "bold",family = "Times"), #set the size, boldness, and style of the legend title
          legend.text = element_text(size = 12,family = "Times"), #set the size and style of the legend text
          legend.position = c(0.9,0.8))+ #specify the position of legend within the plot
    coord_cartesian(ylim = c(0, 15))+ #set the limit of y axis to the range of 0 to 15
    scale_y_continuous(breaks = seq(from = 0, to = 15, by = 2), #set a tick mark on the y-axis every 2 unit increase in y
                       labels = seq(from = 0, to = 15, by = 2)) #label the tick on the y-axis every 2 unit increase in y

ggsave("Papervis.png", dpi=300, dev='png', height=5, width=7, units="in") #save the graph
```

### For Poster

```{r}
Clean_AItrust %>%
    ggplot(aes(y = TrustDiff, x = FramingM, color = Accuracy, width = 20, height = 50))+
    stat_summary(
        fun = mean, 
        geom = "point",
        size = 4) +
    stat_summary(
        fun = mean, 
        geom = "line", 
        size = 1, 
        aes(group = Accuracy)) +
    stat_summary(
        fun.data = mean_se, 
        geom = "errorbar", 
        width = 0.2, 
        size = 1)+
    labs(x = "Framing Message", 
         y = "Change in Trust in ChatGPT",
         title = str_wrap("Do Negative Framing Message and ChatGPT Accurary influence College Students' Trust in ChatGPT?", width = 50)) +
    theme(
          plot.title = element_text(size = 30, face = "bold", hjust = 0, margin = margin(b = 15)), #Set the size, boldness and margin of the plot title, and left align the title
          axis.title.x  = element_text(size = 27, face = "bold", margin = margin(t = 15)),#set the size, boldness, and top margin of the x axis title 
          axis.title.y  = element_text(size = 27, face = "bold",margin = margin(r = 10)),#set the size, boldness, and right marginof the y axis title 
          axis.text = element_text(size = 27, color = "black"), #set the size, color, and style of both axis texts
          legend.title = element_text(size = 27, face = "bold"), #set the size, boldness, and style of the legend title
          legend.text = element_text(size = 27) #set the size and style of the legend text
          )+
    coord_cartesian(ylim = c(0, 15))+#set the limit of y axis to the range of 0 to 15
    scale_y_continuous(breaks = seq(from = 0, to = 15, by = 2), #set a tick mark on the y-axis every 2 unit increase in y
                       labels = seq(from = 0, to = 15, by = 2))#label the tick on the y-axis every 2 unit increase in y

ggsave("Postervis.png", dpi=300, dev='png', height=10, width = 14, units="in") #save the graph
```

# Data Analysis

### Run 2\*2 Between Subject Factorial model

Test for:

1.  Main effect of Factor: Framing Message
2.  Main effect of Factor: ChatGPT Accuracy
3.  Interaction between two factors

```{r}
TrustDiff_model <- lm(TrustDiff ~ FramingM * Accuracy, data= Clean_AItrust) 
TrustDiff_model
```

```{r}
anova(TrustDiff_model)
summary(TrustDiff_model)
```

### Pairwise Comparisons

Means for each condition

```{r}
emmeans(TrustDiff_model,"FramingM")
emmeans(TrustDiff_model,"Accuracy")
emmeans(TrustDiff_model,"FramingM","Accuracy") 
```

Tukey's Result

```{r}
TukeyHSD(TrustDiff_model)
```

### Number of participants in each condition

```{r}
Clean_AItrust %>% 
  group_by(FramingM,Accuracy) %>% 
  summarize( ParticipantsNum = n(),
             meantd = mean(TrustDiff),
             sdtrustdiff = sd(TrustDiff))
```

### Demographic: Participants' Academic Standing

```{r}
tally(Clean_AItrust$College) 
```

# Exploratory Analysis (not included in paper)

### Trust3Mean Visualization

```{r}
Clean_AItrust %>%
    ggplot(aes(y = Trust3Mean, x = FramingM, color = Accuracy))+
    stat_summary(
        fun = mean, 
        geom = "point", 
        size = 4) +
    stat_summary(
        fun = mean, 
        geom = "line", 
        size = 1, 
        aes(group = Accuracy)) +
    stat_summary(
        fun.data = mean_se, 
        geom = "errorbar", 
        width = 0.2, 
        size = 1)+
    labs(x = "Framing Messaging", 
         y = "Trust towards three questions of ChatGPT",
         ) +
    theme(text = element_text(size = 16))
```

#### Run 2\*2 Between Subject Factorial model on Trust3mean

```{r}
Trust3Mean_Accuracy_model <- lm(Trust3Mean ~ Accuracy + FramingM, data = Clean_AItrust)
Trust3Mean_Accuracy_model 
```

```{r}
anova(Trust3Mean_Accuracy_model)
```

### Feeling3Mean Visualization

```{r}
Clean_AItrust %>%
    ggplot(aes(y =Feeling3Mean , x = FramingM, color = Accuracy))+
    stat_summary(
        fun = mean, 
        geom = "point", 
        size = 4) +
    stat_summary(
        fun = mean, 
        geom = "line", 
        size = 1, 
        aes(group = Accuracy)) +
    stat_summary(
        fun.data = mean_se, 
        geom = "errorbar", 
        width = 0.2, 
        size = 1)+
    labs(x = "Framing Message", 
         y = "Feeling towards ChatGPT",
         title = "Effect of Framing Message and ChatGPT Accurary on 
    College Students' General Emotion towards ChatGPT") +
    theme(text = element_text(size = 15))

```

#### Run 2\*2 Between Subject Factorial model on Feeling3mean

```{r}
Feeling_model <- lm(Feeling3Mean ~ FramingM * Accuracy, data = Clean_AItrust)
Feeling_model
```

```{r}
anova(Feeling_model)
```

### Feeling_Trust visualization

```{r}
Clean_AItrust %>%
    ggplot(aes(y = Feeling3Mean , x = Trust3Mean))+
  geom_point()+
  geom_smooth(method = "lm")+
    labs(x = "Trust in ChatGPT", 
         y = "Feeling in ChatGPT",
         title = "Correlation between Trust and Feeling in ChatGPT") +
    theme(text = element_text(size = 15))
```

#### Run Simple Linear Regression Model

```{r}
lm(Trust3Mean ~ Feeling3Mean, data = Clean_AItrust)
```
